---
title: "Starting R for Data Analysis"
author: "Amy Goldlist"
date: "05/05/2022"
output: html_fragment
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)



#output: md_document:
 # variant:  markdown_phpextra+backtick_code_blocks

```



Here's some fun information for you to do on your own:


<iframe width="560" height="315" src="https://www.youtube.com/embed/wBybv2ikk7U" title="What this si about" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Installing R on your own computer:

<iframe width="560" height="315" src="https://www.youtube.com/embed/RZxT3UrshsQ" title="Installing R" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


### Links:

There are two things to install.  RStudio is the IDE (development environment), and R is the scripting language.  They are both free!

* <https://rstudio.com/products/rstudio/download/>
* <https://cran.rstudio.com/>


## R as a giant calculator:
```{r}

## when I start a line with a #, it's a comment

### Try to understand how the code is working!


# The <- (or alt -)  assigns a variable
x <- 5
## now x is always 5


##  what is 8 times 5?
8*x


## now make x be 7
x <- 7

## hmm, 8x is different!
8*x

```


## A data problem:

See the "GiG" worksheet, which is included in this bundle.  This is from the textbook "Business Analytics: Communicating with Numbers" by Jaggia et al, available from McGraw Hill.

This data set contains Employees, categorized by Wage, Industry and Job.  It contains missing info.  Here's a glimpse:

| **EmployeeID** | **HourlyWage** | **Industry**     | **Job**        |
|------------|------------|--------------|------------|
| 20         | 26.09      | Construction | Consultant |
| 21         | 49.59      | Construction |            |
| 22         | 47.97      | Construction | Accountant |
| 23         | 48.77      | Construction | Engineer   |
| 24         | 42.58      |              | Sales Rep  |
| 25         | 49.7       | Automotive   | Engineer   |


### Questions

We are now going to answer the following questions:

##### Find number of missing Values:
* Hourly Wage
* Industry
* Job 

##### The Number of employees who:

* work in the automotive industry
* Earn More than $30 per hour
* Automotive Industry and earn more than $30 per hour

##### Find the Hourly wages:
* Lowest: 
* Highest:
* Lowest accountant in automotive: 
* Highest accountant in automotive: 
* Lowest accountant in tech: 
* Highest accountant in tech: 

## Let's try Excel

<iframe width="560" height="315" src="https://www.youtube.com/embed/J80i6S7fUIs" title="It's Excel!" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


## And R:
<iframe width="560" height="315" src="https://www.youtube.com/embed/PUDEHnBgUnw" title="And finally R" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

### Packages (or libraries)

Why recreate the wheel, when someone has already doen the work for us?

```{r }
### if this is your first time using this, open these files:
# Dlete the # at the beginign of the line!

## Go to the end of each line and press ctrl+ enter
# install.packages("dplyr")
# install.packages("openxlsx")
# install.packages("ggplot2")
# install.packages("palmerpenguins")

## opening up these libraries or packages lets us use them
library(dplyr)  #dplyr is really great for organizing dataframes
library(openxlsx)  #openxlsx lets us read and write to Excel files. 
library(ggplot2) ## this is for making visualizations
library(palmerpenguins) # this is actually just cute penguins!

```


### Open the file
```{r}
## Wait, where am I on my computer?

getwd()

## Let's read in our Excel file.  The sheet option tells us which worksheet to use
gig <- read.xlsx("jaggia_ba_1e_ch02_Data_Files.xlsx",sheet ="Gig")

```


### And let's look at the file:

```{r}
## how big is our data set?
dim(gig)  

## look at the first few rows
head(gig)  

## look at the whole thing in a different window
# View(gig)  


## what are the columns like?  str = structure
str(gig)
```


### Answering the questions


##### Find number of missing Values:
* Hourly Wage

```{r}
## ctrl shift m makes the cool "pipe" %>%
##step1:  pull up the gig dataset
gig %>% 
  ##step 2:  filter only the blank hourly wage
  filter(is.na(HourlyWage))

```

* Industry

```{r}
gig %>% 
  ## filter the blanks in Industry
  filter(is.na(Industry))
```

* Job

```{r}
### and the blank jobs
gig %>% 
  filter(is.na(Job))
```


##### The Number of employees who:

* work in the automotive industry
```{r}
## Now lets try to count stuff using summarize...

## first take our whole data.frame
gig %>% 
  ## group by industry
  group_by(Industry) %>%
  #then count the numbers n()
  summarize(n())
```


* Earn More than $30 per hour

```{r}
gig %>% 
  ## filter by wage greater than $30
  filter(HourlyWage> 30) %>% 
  ##and count them
  count()

```

* Automotive Industry and earn more than $30 per hour

```{r}
gig %>% 
  filter(HourlyWage> 30) %>% 
  ## group by industry
  group_by(Industry) %>%
  #then count the numbers n()
  summarize(n())
```


##### Find the Hourly wages:
* Lowest and Highest:

```{r}
gig %>% 
  #then find maximum and minimum HourlyWage
  summarize(min(HourlyWage),max(HourlyWage))
```


* Lowest and Highest accountant in automotive / Tech : 

```{r}
## Step 1: pull up gig dataset
gig %>% 
  ## Step 2: filter by only Accountants
  filter(Job == "Accountant") %>% 
  ## Step 3: group by industry
  group_by(Industry) %>% 
  #then find maximum and minimum wage
  summarize(min(HourlyWage),max(HourlyWage))

```


###  Pretty pictures

Just some basic plots:

```{r}

## try commenting and uncommenting:

#plot(gig)


# gig %>% ggplot(aes(y =HourlyWage))+
#   geom_boxplot()
# 
# gig %>% ggplot(aes(y =HourlyWage, colour = Industry))+
#   geom_boxplot()+
#   theme_bw
# 
# gig %>% ggplot(aes(y =HourlyWage, colour = Job))+
#   geom_boxplot()+
#   theme_bw()
# 
# 
# 
# gig %>% ggplot(aes(x =HourlyWage))+
#   geom_histogram()+
#   theme_bw()
# 
 gig %>% ggplot(aes(x =HourlyWage))+
   geom_histogram(aes(fill = Industry))+
   theme_bw()
```
 
## GEtting Fancy!

Let's look at a different dataset.  This one is all about penguins.  Intrigued?  (<https://allisonhorst.github.io/palmerpenguins/>)

```{r}


## Look at the data



## Look at penguins

penguins


```

### References:

* Take some time and learn about R from the experts: <https://education.rstudio.com/learn/beginner/>
* This is an hour long webinar, but will give you a good overview of data wrangling in R: <https://rstudio.com/resources/webinars/a-gentle-introduction-to-tidy-statistics-in-r/>
* Follow some tutorials at <https://stat545.com/>
* Like this example?  Buy the textbook: <https://www.mheducation.com/highered/product/business-analytics-jaggia-kelly/M9781260785005.html>

