---
title: "Fun Soccer Worksheet"
author: "Amy Goldlist"
date: "08/04/2022"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  ## this is coding instructions, ignore!


```


## Open the libraries we will need:

As well as get that fun data!

```{r}
library(dplyr) ## dplyr is for working with data frames 
library(openxlsx) ## lets us liaise with Excel
library(stringr) ## Extract words to make it faster!



## read in the data
df <- read.xlsx("soccer.xlsx", sheet = "Detailorders", startRow = 2)


```


## Basic Data

First, let's look at what we have:
```{r }


df
```

This we can look at the structure:

```{r }


df %>% str()
```


## Reshape the data:

We want the data to look like this:


| Player        | Item   |Size |Received
|--------|-----------|------------|------------|
| 1 |     Jersey|YL |Yes|
| 7    | Socks | Med    |No|


Let's group the data:



```{r, echo = TRUE}

## make a string of names
col_names <- colnames(df)

## WE want to see just the clothing:
interm_list <- col_names[2:27]

interm_list

size_list <- c()
rec_list <- c()
for (i in 1:13){
  size_list <- c(size_list,interm_list[2*i-1])
  rec_list <- c(rec_list,interm_list[2*i])
}

item_list <- str_remove(rec_list,".Received")


## Make a blank df

Soccer_df <- data.frame("Player"=numeric(), "Item"= character(), "Size"=character(), "Recieved"=character())



for (i in 1:13){


  newdf <- df %>% 
    select(Player, size_list[i], rec_list[i])
  
  names(newdf) <- c("Player", "Size", "Received")
  newdf$Item <- item_list[i]
  
  newdf <- newdf %>% 
    select(Player, Item, Size, Received)
  
Soccer_df <- rbind(Soccer_df, newdf)
  
}


## Inspect the first 5 rows:

Soccer_df %>% head()


```


Now we will look at grouping the data.

I filtered out anything where the size was "NA"


```{r , echo=TRUE}


summary <- Soccer_df %>% 
  filter(!is.na(Size)) %>% 
  group_by(Item, Size, Received) %>% 
  summarize(Count = n())

to_save <- Soccer_df %>% 
  filter(!is.na(Size)) %>% 
  group_by(Item, Size, Received) %>% 
  summarize(Count = n()) %>% 
  filter(Received == "No")

to_save

```

## Save

Finally, we cna save that file as a new Excel sheet - or you could mkae it an old Excel Sheet, whatever!

```{r}
file_name <- paste0("Soccer_",Sys.Date(),".xlsx")
file_name


wb <- createWorkbook("Soccer_Gear")

addWorksheet(wb, sheetName = "original")
writeData(wb, sheet = "original", x = df)

addWorksheet(wb, sheetName = "Not_Received")
writeData(wb, sheet = "Not_Received", x = to_save)

addWorksheet(wb, sheetName = "Summary")
writeData(wb, sheet = "Summary", x = summary)



saveWorkbook(wb, file_name, overwrite = TRUE)



```


